Progress Report

### 2016-01-01 Report Dec 2015/Report.tex
SENIA-1 simulation results
Neumann series expansion refined by 1-time Newton Iteration

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
The abstract goes here.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
IEEEtran, journal, \LaTeX, paper, template.
\end{IEEEkeywords}

@@

Bit sequences, which are modulated to complex symbols, are transmitted by the users over a flat fading channel. The discrete time model of the system is given by:
\begin{equation}
\mathbf{y}=\mathbf{H}\mathbf{s}+\mathbf{n},   \label{discrete time MIMO system}
\end{equation}
where $\mathbf{y}\in\mathbb{C}^{N_{r}\times 1}$ is the received symbol vector, $\mathbf{s}\in \mathbb{C}^{N_{t}\times 1}$ is the transmitted symbol vector, with components that are mutually independent and taken from a finite signal constellation alphabet $\mathbb{O}$ (e.g., BPSK, 4-QAM, 16-QAM, 64-QAM), $|\mathbb{O}|=M$. The transmitted symbol vectors $\mathbf{s}\in \mathbb{O}^{N_{t}}$, satisfy $\mathbb{E}[\mathbf{s}\mathbf{s}^{H}]=\mathbf{I}_{N_t}E_{s}$, where $E_{s}$ denotes the symbol average energy, $\mathbb{E}[\cdot]$ denotes the expectation operation, $\mathbf{I}_{N_{t}}$ denotes identity matrix of size $N_{t}\times N_{t}$. Furthermore $\mathbf{H}\in \mathbb{C}^{N_{r}\times N_{t}}$ denotes the Rayleigh fading channel propagation matrix. $\mathbf{H}_{ij}$ denotes the component of $\mathbf{H}$ at $i$th row and $j$th column, representing the channel response from $i$th receive antenna to the $j$th transmit antenna. Each component is independent identically distributed (i.i.d) circularly symmetric complex Gaussian (CSCG) random variable with unit variance, denoted by $\mathbf{H}_{ij}\sim \mathbb{C}N(0,1)$. Finally, $\mathbf{n}\in \mathbb{C}^{N_{r}}$ is the additive white Gaussian noise (AWGN) vector with zero mean components and $\mathbb{E}[\mathbf{n}\mathbf{n}^{H}]=\mathbf{I}_{N_{r}}N_{0}$, where $N_{0}$ denotes the noise power spectrum density, and hence $\frac{E_{s}}{N_{0}}$ is the symbol signal to noise ratio (SNR).

@@

Consider an alternative representation of MLD principle, let $\mathbf{H}_{1}\in \mathbb{C}^{N_{r}\times N}$ denotes a sub matrix composed of $N$ columns from $\mathbf{H}$, where $1\leq N\leq N_{t}$, let $\mathbf{s}_{1}\in \mathbb{C}^{N}$ denote the sub symbol vector whose components are transmitted from the users corresponding to $\mathbf{H}_{1}$. Similarly, let $\mathbf{H}_{2}\in \mathbb{C}^{N_{r}\times (N_{t}-N)}$ denotes the sub matrix composed of the remaining columns from $\mathbf{H}$ and $\mathbf{s}_{2}\in \mathbb{C}^{(N_{t}-N)}$ is the sub symbol vector whose components are transmitted by the users corresponding to $\mathbf{H}_{2}$. Thus (\ref{discrete time MIMO system}) can be rewritten as 
\begin{equation}
\mathbf{y}=\mathbf{H}_{1}\mathbf{s}_{1}+\mathbf{H}_{2}\mathbf{s}_{2}+\mathbf{n}.
\label{alternative MIMO system}
\end{equation}  
Let $\hat{\mathbf{s}}_{1}, \hat{\mathbf{s}}_{2}$ denote the estimations of $\mathbf{s}_{1}, \mathbf{s}_{2}$, we have $\hat{\mathbf{s}}_{1}\in [\hat{\mathbf{s}}^{1}_{1}, \hat{\mathbf{s}}^{2}_{1}, \cdots \hat{\mathbf{s}}^{K}_{1}],, K=M^{N}$ and $\hat{\mathbf{s}}_{2}\in [\hat{\mathbf{s}}^{1}_{2}, \hat{\mathbf{s}}^{2}_{2}, \cdots \hat{\mathbf{s}}^{Q}_{2}], Q=M^{N_{t}-N}$, MLD in (\ref{MLD}) can be rewritten as 
\begin{equation}
[\tilde{k}, \tilde{q}]=\arg\min_{k\in [1,2\ldots K]}\min_{q\in [1,2,\ldots, Q]}||\mathbf{y}-\mathbf{H}_{1}\hat{\mathbf{s}}^{k}_{1}-\mathbf{H}_{2}\hat{\mathbf{s}}^{q}_{2}||^{2},\\
\mathbf{s}^{ML}_{1}=\hat{\mathbf{s}}^{\tilde{k}}_{1}, \mathbf{s}^{ML}_{2}=\hat{\mathbf{s}}^{\tilde{q}}_{2},
\label{alternative MLD}
\end{equation}
(\ref{alternative MLD}) can be divided into three steps, first define 
\begin{equation}
\mathbf{y}^{k}=\mathbf{y}-\mathbf{H}_{1}\hat{\mathbf{s}}^{k}_{1}\quad k\in [1,2,\ldots, K],
\label{first step alternative MLD}
\end{equation}
Then solve 
\begin{equation}
\hat{\mathbf{x}}_{2}^{k}=\arg\min_{\hat{\mathbf{s}}_{2}^{q}\in [\hat{\mathbf{s}}_{2}^{1},\hat{\mathbf{s}}_{2}^{3}\ldots, \hat{\mathbf{s}}_{2}^{Q}]}||\mathbf{y}^{k}-\mathbf{H}_{2}\hat{\mathbf{s}}_{2}^{q}||^{2}, 
\label{second step alternative MLD}
\end{equation}
\begin{equation}
\tilde{k}=\arg\min_{k\in [1,2,\ldots, K]}||\mathbf{y}^{k}-\mathbf{H}_{2}\hat{\mathbf{x}}_{2}^{k}||^{2},
\label{third step alternative MLD}
\end{equation}
Finally we have $\mathbf{s}_{1}^{ML}=\hat{\mathbf{s}}_{1}^{\tilde{k}}, \mathbf{s}_{2}^{ML}=\hat{\mathbf{x}}_{2}^{\tilde{k}}=\hat{\mathbf{s}}_{2}^{\tilde{q}}$.
% needed in second column of first page if using \IEEEpubid
%\IEEEpubidadjcol

@@

\section{Minimum Achievable Diversity based Channel Partition}
\subsection{Diversity Maximization Selection}\label{DMS analysis}
Based on the alternative form of MLD in (\ref{alternative MLD})-(\ref{third step alternative MLD}), General Parallel Interference Cancellation (GPIC) algorithm first generate a list of symbol vector candidates and then choose the candidate with the minimum Euclidean distance as the solution. To elaborate further, GPIC detectors first generate a list of symbol vector candidate by exhaustive search for all the possible $\hat{\mathbf{s}}_{1}$ as shown in (\ref{first step alternative MLD}), then rather than performing exhaustive search in (\ref{second step alternative MLD}), GPIC detectors exploit the low complexity sub optimal linear detectors (i.e., zeros forcing (ZF) and minimum mean square error (MMSE)) and their successive interference cancellation (SIC) counterparts to get the estimation of $\mathbf{s}^{k}_{2}$, $k=[1,2,\ldots, K]$. For sake of simplicity, hereinafter we use LD/SIC. Then the best candidate in the list with the minimum Euclidean distance is chosen as the solution.

Performance analysis of GPIC algorithm is provided in \cite{radji2009interference}, here we briefly conclude the main results in \cite{radji2009interference}. At the receiving side, the frame error (detection error) is defined as there is at least one erroneous symbol in the estimation of transmitted symbol vector. Let $P_{e}$ denotes the average frame error probability. The diversity order, denoted by $d$ is defined as the slope of the $P_{e}$ in log-scale in high SNR region, which is given by\cite{zheng2003diversity}
\begin{equation}
d=-\lim_{SNR\to\infty}\frac{\log{P_{e}}}{\log(SNR)},
\label{diversity}
\end{equation}
Let $P_{e}^{ML}$ denotes the average frame error probability of MLD, The average frame error probability of GPIC algorithm $P_{et}=\mathbb{E}_{\mathbf{H}, \mathbf{s}}[Pr(\hat{\mathbf{s}}\neq \mathbf{s}|\mathbf{H}, \mathbf{s})]$ and $P_{e2}=\mathbb{E}_{\mathbf{H}_{2}, \mathbf{s}_{2}}(Pr(\hat{\mathbf{s}}^{k_{1}}_{2}\neq \mathbf{s}_{2}|\mathbf{H}_{2}, \mathbf{s}_{2}))$, where $P_{r}(A)$ denotes the probability that a event A occurs, $\hat{\mathbf{s}}_{1}^{k_{1}}=\mathbf{s}_{1}$, therefore $P_{e2}$ is the average frame error probability of the LD/SICs for the sub system in which the interference from the users that transmitted $\mathbf{s}_{1}$ are perfectly cancelled from the observation $\mathbf{y}$. given by
\begin{equation}
\mathbf{y}^{k_{1}}=\mathbf{y}-\mathbf{H}_{1}\hat{\mathbf{s}}_{1}^{k_{1}},
\label{perfect IC}
\end{equation} 
because $\hat{\mathbf{s}}^{k1}_{1}=\mathbf{s}_{1}$, we have
\begin{equation}
\mathbf{y}^{k1}=\mathbf{H}_{2}\mathbf{s}_{2}+\mathbf{H}_{1}(\mathbf{s}_{1}-\hat{\mathbf{s}}_{1}^{k1})+\mathbf{n}=\mathbf{H}_{2}\mathbf{s}_{2}+\mathbf{n},
\label{sub MIMO system}
\end{equation} 
Based on \cite{radji2009interference}, $P_{et}$ is bounded by 
\begin{equation}
\max(P^{ML}_{e}, P_{e2})\leq P_{et}\leq P^{ML}_{e}+P_{e2}, 
\label{average frame error probability of GPIC}
\end{equation}
in (\ref{average frame error probability of GPIC}), the diversity order of MLD is $N_{r}$. Let $d_{t}$ denote the overall diversity order of GPIC detectors and $d_{2}$ denote the diversity order of LD/SICs with the interferences from $\mathbf{s}_{1}$ perfectly cancelled, based on (\ref{diversity}), we have 
\begin{equation}
\lim_{SNR\to\infty}P^{ML}_{e}\propto SNR^{-N_{r}},
\label{MLD diversity}
\end{equation}
\begin{equation}
\lim_{SNR\to\infty}P_{et}\propto SNR^{-d_{t}},
\label{GPIC diversity}
\end{equation}
\begin{equation}
\lim_{SNR\to\infty}P_{e2}\propto SNR^{-d_{2}},
\label{LD diversity}
\end{equation}
 now consider the following conditions
\begin{itemize}
\item $d_{2}>N_{r}$,
\begin{equation}
\lim_{SNR\to\infty}\frac{P_{e2}}{P^{ML}_{e}}=SNR^{N_{r}-d_{2}}=0,
\label{condition1sub1}
\end{equation}
based on (\ref{average frame error probability of GPIC}), we have
\begin{equation}
\lim_{SNR\to\infty}\frac{P_{et}}{P^{ML}_{e}}=1,
\label{condition1sub2}
\end{equation}
thus $d_{t}=N_{r}$.
\item $d_2=N_{r}$,
we have 
\begin{equation}
\lim_{SNR\to\infty}P_{et}\propto SNR^{-N_{r}},
\label{condition2sub1}
\end{equation}
thus $d_{t}=N_{r}$.
\item $d_{2}<N_{r}$,
\begin{equation}
\lim_{SNR\to\infty}\frac{P^{ML}_{e}}{P_{e2}}=SNR^{d_{2}-N_{r}}=0,
\label{condition3sub1}
\end{equation}
based on (\ref{average frame error probability of GPIC}), we have  
\begin{equation}
\lim_{SNR\to\infty}P_{et}=P_{e2}\propto SNR^{-d_{2}},
\label{condition3sub2}
\end{equation}
thus 
$d_{t}=d_{2}$.
\end{itemize} 
Therefore when $d_{2}\geq N_{r}$, GPIC algorithm can achieve ML performance asymptotically, when $d_{2}<N_{r}$, the diversity order that GPIC can achieve at high SNR region is equal to that of LD/SICs. In \cite{radji2009interference}, the authors employ diversity maximization selection (DMS) scheme for channel partition, which is originally proposed in \cite{zhang2006diversity}. The optimal diversity order can be guaranteed with low complexity for conventional small MIMO systems.

To elaborate further, for a given number of antennas chosen at the channel partition stage $N$, there are $N_{u}={N_{t}\choose N}$ possible combinations of $[\mathbf{H}_{1}, \mathbf{H}_{2}]$. According to DMS principle, the maximum diversity order of LD/SIC is achieved by choosing the subset $\mathbf{H}_{2}$ that has the strongest weakest substream, in a sense of post processing SNR. Therefore, when using LD/SIC detectors as the sub optimal detector, the subset $\mathbf{H}^{opt}_{2}$ selected based on DMS principle should be\cite{zhang2006diversity} 
\begin{IEEEeqnarray}[\relax]{l}
\mathbf{H}^{opt}_{2}=\mathbf{H}^{p}_{2},\\
p=\arg\min_{j=1,2,\ldots, N_{u}}\theta_{j},\\
\theta_{j}=\max_{k=1,2,\ldots, N_{t}-N}((\mathbf{H}_{2}^{j})^{H}\mathbf{H}_{2}^{j}+SNR^{-1}\mathbf{I})_{kk}^{-1},\quad \mathbf{H}_{2}^{j}\in[\mathbf{H}_{2}^{1}, \mathbf{H}_{2}^{2}, \ldots, \mathbf{H}_{2}^{N_{u}}];
\label{DMS principle}
\end{IEEEeqnarray}   
Where $\mathbf{A}_{kk}$ denotes the $k$th diagonal component of matrix $\mathbf{A}$. The diversity order of LD/SIC detector after DMS channel partition process is given by 
$d_{2DMS}=(N+1)(N_{r}-N_{t}+N+1)$. In \cite{radji2009interference}, the authors derive the minimum number of antennas $N_{min}$ at channel partition stage that can guarantee $d_{2DMS}\geq N_{r}$, which is given by 
\begin{equation}
N_{min}=\lceil \sqrt{\frac{(N_{r}-N_{t})^{2}}{4}+N_{r}}-\frac{N_{r}-N_{t}}{2}-1\rceil,
\label{Nmin DMS}
\end{equation}
where $\lceil\alpha\rceil$ denotes the minimum integer no less than $\alpha$.
\subsection{Minimum Achievable Diversity: How many transmit antennas do we need to chose in LS-MIMO?}
In LS-MIMO V-BLAS systems, where the substreams are transmitted independently, the maximum diversity order is $N_{r}$, which is extremely large under the condition that there are tens to hundreds of receive antennas. On the one hand, diversity order is achieved at high SNR region, however, in practical, such a high diversity order is not necessary, because the SNR region of interest is the range in which the bit error rate (BER) or symbol error rate (SER) can be $10^{-5}$ to $10^{-7}$. On the other hand, the computational complexity to guarantee ML performance by DMS channel partition is excessive in LS-MIMO, for example, when $N_{r}=N_{t}$, based on (\ref{Nmin DMS}), the $N_{min}=\lceil\sqrt{N_{r}}-1\rceil$, the number of the symbol vector candidates in the list is $M^{N_{min}}$.

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
This report consists of two major parts. In the first part, based on the performance analysis of list based algorithm, minimum achievable diversity (MAD) channel selection principle for Large-Scale MIMO (LS-MIMO) systems is presented. MAD principle can reduce the detection complexity by sacrificing redundant asymptotic diversity gain in LS-MIMO. Computer simulation results for the performance of MAD principle are presented for different MIMO system configurations. In the second part, theoretical analysis are provided for channel hardening phenomenon, including logarithmic expectation and probability distribution of orthogonality measure, which is defined to evaluate the orthogonality of the channel matrix. 
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
%\begin{IEEEkeywords}
%IEEEtran, journal, \LaTeX, paper, template.
%\end{IEEEkeywords}

@@

Bit sequences, which are modulated to complex symbols, are transmitted by the users over a flat fading channel. The discrete time model of the system is given by:
\begin{equation}
\mathbf{y}=\mathbf{H}\mathbf{s}+\mathbf{n},   \label{discrete time MIMO system}
\end{equation}
where $\mathbf{y}\in\mathbb{C}^{N_{r}\times 1}$ is the received symbol vector, $\mathbf{s}\in \mathbb{C}^{N_{t}\times 1}$ is the transmitted symbol vector, with components that are mutually independent and taken from a finite signal constellation alphabet $\mathbb{O}$ (e.g., BPSK, 4-QAM, 16-QAM, 64-QAM), $|\mathbb{O}|=M$. The transmitted symbol vectors $\mathbf{s}\in \mathbb{O}^{N_{t}}$, satisfy $\mathbb{E}[\mathbf{s}\mathbf{s}^{H}]=\mathbf{I}_{N_t}E_{s}$, where $E_{s}$ denotes the symbol average energy, $\mathbb{E}[\cdot]$ denotes the expectation operation, $\mathbf{I}_{N_{t}}$ denotes identity matrix of size $N_{t}\times N_{t}$. Furthermore $\mathbf{H}\in \mathbb{C}^{N_{r}\times N_{t}}$ denotes the Rayleigh fading channel propagation matrix. $\mathbf{H}_{ij}$ denotes the component of $\mathbf{H}$ at $i$th row and $j$th column, representing the channel response from $i$th receive antenna to the $j$th transmit antenna. Each component is independent identically distributed (i.i.d) circularly symmetric complex Gaussian (CSCG) random variable with unit variance, denoted by $\mathbf{H}_{ij}\sim \mathbb{C}N(0,1)$. Finally, $\mathbf{n}\in \mathbb{C}^{N_{r}\times 1}$ is the additive white Gaussian noise (AWGN) vector with zero mean components and $\mathbb{E}[\mathbf{n}\mathbf{n}^{H}]=\mathbf{I}_{N_{r}}N_{0}$, where $N_{0}$ denotes the noise power spectrum density, and hence $\frac{E_{s}}{N_{0}}$ is the symbol signal to noise ratio (SNR).

@@

Consider an alternative representation of MLD principle, let $\mathbf{H}_{1}\in \mathbb{C}^{N_{r}\times N}$ denotes a sub matrix composed of $N$ columns from $\mathbf{H}$, where $1\leq N\leq N_{t}$, let $\mathbf{s}_{1}\in \mathbb{C}^{N\times 1}$ denote the sub symbol vector whose components are transmitted from the users corresponding to $\mathbf{H}_{1}$. Similarly, let $\mathbf{H}_{2}\in \mathbb{C}^{N_{r}\times (N_{t}-N)}$ denotes the sub matrix composed of the remaining columns from $\mathbf{H}$ and $\mathbf{s}_{2}\in \mathbb{C}^{(N_{t}-N)\times 1}$ is the sub symbol vector whose components are transmitted by the users corresponding to $\mathbf{H}_{2}$. Thus (\ref{discrete time MIMO system}) can be rewritten as 
\begin{equation}
\mathbf{y}=\mathbf{H}_{1}\mathbf{s}_{1}+\mathbf{H}_{2}\mathbf{s}_{2}+\mathbf{n}.
\label{alternative MIMO system}
\end{equation}  
Let $\hat{\mathbf{s}}_{1}, \hat{\mathbf{s}}_{2}$ denote the estimations of $\mathbf{s}_{1}, \mathbf{s}_{2}$, we have $\hat{\mathbf{s}}_{1}\in [\hat{\mathbf{s}}^{1}_{1}, \hat{\mathbf{s}}^{2}_{1}, \cdots \hat{\mathbf{s}}^{K}_{1}], K=M^{N}$ and $\hat{\mathbf{s}}_{2}\in [\hat{\mathbf{s}}^{1}_{2}, \hat{\mathbf{s}}^{2}_{2}, \cdots \hat{\mathbf{s}}^{Q}_{2}], Q=M^{N_{t}-N}$, then MLD in (\ref{MLD}) can be rewritten as 
\begin{equation}
[\tilde{k}, \tilde{q}]=\arg\min_{k\in [1,2\ldots K]}\min_{q\in [1,2,\ldots, Q]}||\mathbf{y}-\mathbf{H}_{1}\hat{\mathbf{s}}^{k}_{1}-\mathbf{H}_{2}\hat{\mathbf{s}}^{q}_{2}||^{2},\\
\mathbf{s}^{ML}_{1}=\hat{\mathbf{s}}^{\tilde{k}}_{1}, \mathbf{s}^{ML}_{2}=\hat{\mathbf{s}}^{\tilde{q}}_{2},
\label{alternative MLD}
\end{equation}
(\ref{alternative MLD}) can be divided into three steps, first define 
\begin{equation}
\mathbf{y}^{k}=\mathbf{y}-\mathbf{H}_{1}\hat{\mathbf{s}}^{k}_{1}\quad k\in [1,2,\ldots, K],
\label{first step alternative MLD}
\end{equation}
Then solve 
\begin{equation}
\hat{\mathbf{x}}_{2}^{k}=\arg\min_{\hat{\mathbf{s}}_{2}^{q}\in [\hat{\mathbf{s}}_{2}^{1},\hat{\mathbf{s}}_{2}^{3}\ldots, \hat{\mathbf{s}}_{2}^{Q}]}||\mathbf{y}^{k}-\mathbf{H}_{2}\hat{\mathbf{s}}_{2}^{q}||^{2}, 
\label{second step alternative MLD}
\end{equation}
\begin{equation}
\tilde{k}=\arg\min_{k\in [1,2,\ldots, K]}||\mathbf{y}^{k}-\mathbf{H}_{2}\hat{\mathbf{x}}_{2}^{k}||^{2},
\label{third step alternative MLD}
\end{equation}
Finally we have $\mathbf{s}_{1}^{ML}=\hat{\mathbf{s}}_{1}^{\tilde{k}}, \mathbf{s}_{2}^{ML}=\hat{\mathbf{x}}_{2}^{\tilde{k}}=\hat{\mathbf{s}}_{2}^{\tilde{q}}$.
% needed in second column of first page if using \IEEEpubid
%\IEEEpubidadjcol

@@

\section{Minimum Achievable Diversity based Channel Partition}
\subsection{Diversity Maximization Selection}\label{DMS analysis}
Based on the alternative form of MLD in (\ref{alternative MLD})-(\ref{third step alternative MLD}), General Parallel Interference Cancellation (GPIC) algorithm first generate a list of symbol vector candidates and then choose the candidate with the minimum Euclidean distance as the solution. To elaborate further, GPIC detectors first obtain $\mathbf{H}_{2}$ by a certain channel partition rule, then generate a list of symbol vector candidate by exhaustive search for all the possible $\hat{\mathbf{s}}_{1}$ as shown in (\ref{first step alternative MLD}), then rather than performing exhaustive search in (\ref{second step alternative MLD}), GPIC detectors exploit the low complexity sub optimal linear detectors (i.e., zeros forcing (ZF) and minimum mean square error (MMSE)) and their successive interference cancellation (SIC) counterparts to get the estimation of $\mathbf{s}^{k}_{2}$, $k=[1,2,\ldots, K]$. For sake of simplicity, hereinafter we use LD/SIC. Finally the best candidate in the list with the minimum Euclidean distance is chosen as the solution.

Performance analysis of GPIC algorithm is provided in \cite{radji2009interference}, here we briefly conclude the main results in \cite{radji2009interference}. At the receive side, the frame error (detection error) is defined as there is at least one erroneous symbol in the estimation of transmitted symbol vector. Let $P_{e}$ denotes the average frame error probability. The diversity order, denoted by $d$ is defined as the slope of the $P_{e}$ in log-scale in high SNR region, which is given by\cite{zheng2003diversity}
\begin{equation}
d=-\lim_{SNR\to\infty}\frac{\log{P_{e}}}{\log(SNR)},
\label{diversity}
\end{equation}
Let $P_{e}^{ML}$ denotes the average frame error probability of MLD. Define the average frame error probability of GPIC algorithm $P_{et}=\mathbb{E}_{\mathbf{H}, \mathbf{s}}[Pr(\hat{\mathbf{s}}\neq \mathbf{s}|\mathbf{H}, \mathbf{s})]$ and $P_{e2}=\mathbb{E}_{\mathbf{H}_{2}, \mathbf{s}_{2}}(Pr(\hat{\mathbf{s}}^{k_{1}}_{2}\neq \mathbf{s}_{2}|\mathbf{H}_{2}, \mathbf{s}_{2}))$, where $P_{r}(A)$ denotes the probability that a event A occurs. $\hat{\mathbf{s}}_{1}^{k_{1}}=\mathbf{s}_{1}$, therefore $P_{e2}$ is the average frame error probability of the LD/SICs for the sub system $[\mathbf{H}_{2}, \mathbf{s}_{2}]$ in which the interference from the users that transmitted $\mathbf{s}_{1}$ are perfectly cancelled from the observation $\mathbf{y}$. given by
\begin{equation}
\mathbf{y}^{k_{1}}=\mathbf{y}-\mathbf{H}_{1}\hat{\mathbf{s}}_{1}^{k_{1}},
\label{perfect IC}
\end{equation} 
because $\hat{\mathbf{s}}^{k1}_{1}=\mathbf{s}_{1}$, we have
\begin{equation}
\mathbf{y}^{k1}=\mathbf{H}_{2}\mathbf{s}_{2}+\mathbf{H}_{1}(\mathbf{s}_{1}-\hat{\mathbf{s}}_{1}^{k1})+\mathbf{n}=\mathbf{H}_{2}\mathbf{s}_{2}+\mathbf{n},
\label{sub MIMO system}
\end{equation} 
Based on \cite{radji2009interference}, $P_{et}$ is bounded by 
\begin{equation}
\max(P^{ML}_{e}, P_{e2})\leq P_{et}\leq P^{ML}_{e}+P_{e2}, 
\label{average frame error probability of GPIC}
\end{equation}
in (\ref{average frame error probability of GPIC}), the diversity order of MLD is $N_{r}$. Let $d_{t}$ denote the overall diversity order of GPIC detectors and $d_{2}$ denote the diversity order of LD/SICs with the interferences from $\mathbf{s}_{1}$ perfectly cancelled, based on (\ref{diversity}), we have 
\begin{equation}
\lim_{SNR\to\infty}P^{ML}_{e}\propto SNR^{-N_{r}},
\label{MLD diversity}
\end{equation}
\begin{equation}
\lim_{SNR\to\infty}P_{et}\propto SNR^{-d_{t}},
\label{GPIC diversity}
\end{equation}
\begin{equation}
\lim_{SNR\to\infty}P_{e2}\propto SNR^{-d_{2}},
\label{LD diversity}
\end{equation}
 now consider the following conditions
\begin{itemize}
\item $d_{2}>N_{r}$,
\begin{equation}
\lim_{SNR\to\infty}\frac{P_{e2}}{P^{ML}_{e}}=SNR^{N_{r}-d_{2}}=0,
\label{condition1sub1}
\end{equation}
based on (\ref{average frame error probability of GPIC}), we have
\begin{equation}
\lim_{SNR\to\infty}\frac{P_{et}}{P^{ML}_{e}}=1,
\label{condition1sub2}
\end{equation}
thus $d_{t}=N_{r}$.
\item $d_2=N_{r}$,
we have 
\begin{equation}
\lim_{SNR\to\infty}P_{et}\propto SNR^{-N_{r}},
\label{condition2sub1}
\end{equation}
thus $d_{t}=N_{r}$.
\item $d_{2}<N_{r}$,
\begin{equation}
\lim_{SNR\to\infty}\frac{P^{ML}_{e}}{P_{e2}}=SNR^{d_{2}-N_{r}}=0,
\label{condition3sub1}
\end{equation}
based on (\ref{average frame error probability of GPIC}), we have  
\begin{equation}
\lim_{SNR\to\infty}P_{et}=P_{e2}\propto SNR^{-d_{2}},
\label{condition3sub2}
\end{equation}
thus 
$d_{t}=d_{2}$.
\end{itemize} 
Therefore when $d_{2}\geq N_{r}$, GPIC algorithm can achieve ML performance asymptotically, when $d_{2}<N_{r}$, the diversity order that GPIC can achieve at high SNR region is equal to that of LD/SICs. In \cite{radji2009interference}, the authors employ diversity maximization selection (DMS) scheme for channel partition, which is originally proposed in \cite{zhang2006diversity}. The optimal diversity order can be guaranteed with low complexity for conventional small MIMO systems.

To elaborate further, for a given number of antennas chosen at the channel partition stage $N$, there are $N_{u}={N_{t}\choose N}$ possible combinations of $[\mathbf{H}_{1}, \mathbf{H}_{2}]$. According to DMS principle, the maximum diversity order of LD/SIC is achieved by choosing the subset $\mathbf{H}_{2}$ that has the strongest weakest substream, in a sense of post processing SNR. Therefore, when using LD/SIC detectors as the sub optimal detector, the subset $\mathbf{H}^{opt}_{2}$ selected based on DMS principle should be\cite{zhang2006diversity} 
\begin{IEEEeqnarray}[\relax]{l}
\mathbf{H}^{opt}_{2}=\mathbf{H}^{p}_{2},\\
p=\arg\min_{j=1,2,\ldots, N_{u}}\theta_{j},\\
\theta_{j}=\max_{k=1,2,\ldots, N_{t}-N}((\mathbf{H}_{2}^{j})^{H}\mathbf{H}_{2}^{j}+SNR^{-1}\mathbf{I})_{kk}^{-1},\quad \mathbf{H}_{2}^{j}\in[\mathbf{H}_{2}^{1}, \mathbf{H}_{2}^{2}, \ldots, \mathbf{H}_{2}^{N_{u}}];
\label{DMS principle}
\end{IEEEeqnarray}   
Where $\mathbf{A}_{kk}$ denotes the $k$th diagonal component of matrix $\mathbf{A}$. The diversity order of LD/SIC detectors after DMS channel partition process is given by 
$d_{2DMS}=(N+1)(N_{r}-N_{t}+N+1)$\cite{jalden2007maximal}. In \cite{radji2009interference}, the authors derive the minimum number of antennas $N_{min}$ at channel partition stage that can guarantee $d_{2DMS}\geq N_{r}$, which is given by 
\begin{equation}
N_{min}=\lceil \sqrt{\frac{(N_{r}-N_{t})^{2}}{4}+N_{r}}-\frac{N_{r}-N_{t}}{2}-1\rceil,
\label{Nmin DMS}
\end{equation}
where $\lceil\alpha\rceil$ denotes the minimum integer no less than $\alpha$.
\subsection{Minimum Achievable Diversity: How many transmit antennas do we need to chose in LS-MIMO?}
In LS-MIMO V-BLAS systems, where the substreams are transmitted independently, the maximum diversity order is $N_{r}$, which is extremely large under the condition that there are tens to hundreds of receive antennas. On the one hand, diversity order is achieved at high SNR region, however, in practical, such a high diversity order is not necessary, because the SNR region of interest is the range in which the bit error rate (BER) or symbol error rate (SER) can be $10^{-5}$ to $10^{-7}$. On the other hand, the computational complexity to guarantee ML performance by DMS channel partition is excessive in LS-MIMO, for example, when $N_{r}=N_{t}$, based on (\ref{Nmin DMS}), the $N_{min}=\lceil\sqrt{N_{r}}-1\rceil$, the number of the symbol vector candidates in the list is $M^{N_{min}}$.

Therefore, here we consider minimum achievable diversity (MAD) principle, which can reduce the detection complexity by sacrificing redundant asymptotic diversity gain in LS-MIMO. Let $d_{2MAD}$ denote the diversity order achieved by the suboptimal detectors after the MAD based channel partition, $\tilde{N}_{min}$ denote the minimum number of antennas chosen by MAD principle. Our goal is to guarantee the overall diversity is no less than a given minimum diversity while keep $\tilde{N}_{min}$ small. Let $1\leq g<N_{r}$ denote the given minimum achievable diversity order, based on (\ref{condition3sub2}) and the corresponding analysis in section \ref{DMS analysis}, we have
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
 if \quad d_{2MAD}<N_{r}\quad and\quad d_{2MAD}\geq g\\
  d_{t}=d_{2MAD}\geq g.
\end{IEEEeqnarray}
Therefore we can guarantee that the overall diversity order of GPIC algorithm is no less than $g$, in order to derive the minimum number of antennas required based on MAD principle, we define function
\begin{IEEEeqnarray}[\relax]{l}
f(N)=d_{2MAD}-g=(N+1)(N_{r}-N_{t}+N+1)-g,\label{diversity function}
\end{IEEEeqnarray}    
where $N\in [1,2,\ldots, N_{t}]$, our goal is to find minimum $N$, that satisfy $f(N)\geq 0$. The two zero points of quadratic function $f(N)$ are 
\begin{IEEEeqnarray}[\relax]{l}
N_{1}=-\sqrt{\frac{(N_{r}-N_{t})^{2}}{4}+g}-\frac{N_{r}-N_{t}}{2}-1,\\
N_{2}=\sqrt{\frac{(N_{r}-N_{t})^{2}}{4}+g}-\frac{N_{r}-N_{t}}{2}-1
\label{zeros points}
\end{IEEEeqnarray}
when $N\leq N_{1}$ or $N\geq N_{2}$, $f(N)\geq 0$, obviously $N_{1}<0$, there is no feasible $N$ exits that can make the former condition satisfied. Thus, we have 
\begin{equation}
\tilde{N}_{min}=\lceil \sqrt{\frac{(N_{r}-N_{t})^{2}}{4}+g}-\frac{N_{r}-N_{t}}{2}-1\rceil
\label{Nmin MAD}
\end{equation}
\subsection{Computer Simulations}
In this section, computer simulation results for the bit error rate (BER) performances of MAD based GPIC algorithm are presented. As comparisons, Sel-MMSE-OSIC algorithm in \cite{radji2009interference} is also considered, denoted by maximum diversity selection (MDS) scheme. The software testbed is built by C, compiled by GCC compiler version 4.9.2 on 64 bit Debian (release 8.2) Linux systems. The experiments are performed on two desktops, one has Intel I5-4th generation CPU, quad cores, 3.2GHz clock rate and the other one has Intel I7-5th generation CPU, six cores, 3.5GHz clock rate.

We consider uncoded complex spatial multiplexing LS-MIMO systems, For each receive SNR level, the BER are calculated with minimum $10^{5}$ independent channel realizations achieved and minimum 300 symbol errors accumulated. The modulation schemes considered are rectangular M-QAM(4-QAM and 16-QAM) with Gray code labeling. In each channel realization, $N_{t}$ mutually independent randomly generated bit sequences are modulated to complex symbols. $N_{t}$ complex symbols are transmitted over randomly generated Rayleigh flat fading channel, each channel matrix component is CSCG random variable with unit variance. The receive symbol vector is corrupted by AWGN.

For sake of simplicity, we use MAD-$g$ to present the results for minimum achievable diversity selection, where $g$ denotes the minimum achievable diversity. The minimum number of antennas selected by channel partition for a given minimum achievable diversity $g$ is given by (\ref{Nmin MAD}). Maximum diversity selection (MDS) is used to present the Sel-MMSE-OSIC that can achieve ML diversity asymptotically. The minimum number of antennas selected in MDS is given by (\ref{Nmin DMS}).
\begin{table}[htb]
\renewcommand{\arraystretch}{1.3}
\caption{$16\times 16$ system}
\label{t16MIMO4QAM}
\centering
\begin{tabular}{|c|c|c|}
\hline
Scheme&number of antennas selected ($N$)&Asymptotic diversity\\
\hline 
MDS&3&16\\
\hline
MAD-9&2&9\\
\hline
MAD-4&1&4\\
\hline
\end{tabular}
\end{table}
